---
title: 'Final Project EDA'
author: "Seth Metcalf"
output:
  word_document: default
  pdf_document: default
toc: no
---

### Introduction

This is the final project.

##### Importing World Bank Chart
```{r}
library(ggplot2)
worldbank = read.csv("C:\\Users\\sethc\\Documents\\2021 UC Berkeley Spring\\STAT20\\Final Project\\worldbank.csv")
```

*Due 7 May 2021.*

# GDP vs Business Start up Cost
### Code
```{r}
# Variables
GDP = worldbank$GDP # Quantitative, gives a number for the Gross Domestic Product of a country
start_up_cost = worldbank$start_up_cost # Quantitative, gives a percentage of the GNI for the start up of a business

# Graph
ggplot(worldbank,aes(x=GDP, y=start_up_cost)) +
  geom_point(alpha=0.7,color="red") +
  geom_smooth(method="lm",se=F) + 
  ggtitle(label = "GDP vs. Start Up Cost", subtitle = "created by Seth Metcalf") + 
  xlab("GDP") +
  ylab("Start Up Costs")
```

### Explanation

I wanted to see if there was a relation between the start up cost of a business and the GDP of the country. Even though that the graph is slightly skewed due to the outliers in GDP, we can see that there is a negative correlation when trying to compare the two. This can be observed to show that countries that have a higher start up caost for businesses tender to have an overall lower GDP compared to countries that don't have a high start up cost for businesses.



# Amount of Labor vs Amount of Land
### Code
```{r}
# Variables
land = worldbank$land # Quantitative, amount of Land a country has in sq km
labor = worldbank$labor # Quantitative, amount of Labor a country has (people 15 or older, employed or unemployed seeking a job)

# Graph
ggplot(worldbank,aes(x=land, y=labor)) +
  geom_point(alpha=0.7,color="red") +
  geom_smooth(method="lm",se=F) + 
  ggtitle(label = "Land vs. Labor", subtitle = "created by Seth Metcalf") + 
  xlab("Land") +
  ylab("Labor")
```

### Explanation

This graphs the relationship between the amount of labor a country has and the amount of land a country has. The line of best fit that is produced for us is one that shows a positive correlation between the two variables. This means that countries that tend to have more land tend to have a higher population, which makes sense as more land means more resources to expand a nation with. There are a few outliers, but the concentrated group of points in the bottom left of the graph gives us somewhat of an indication that the correlation holds true; it is observed that countries more land tend to have a bigger workforce.
